---
title: "Challenge 7"
author: "Daniel Manning"
description: "Visualization in Multiple Dimensions"
date: "1/16/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_7
---

```{r}
#| label: setup
#| warning: false
#| message: false
library(tidyverse)
library(ggplot2)
library(here)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


## Read in data

```{r}
egg <- here("posts","_data","FAOSTAT_egg_chicken.csv")%>%
  read_csv()
egg_new <- egg %>% 
  select(-c("Domain Code", "Domain", "Area Code", "Element Code", "Item Code", "Year Code", "Flag"))
egg_new
egg_production <- egg_new[egg_new$Element == 'Production',]
egg_production
```

## Time Dependent Visualization 

I chose to use a scatterplot to display the production value over time because these are numeric variables and the scatterplot can be used to reveal relationships between them. In addition, I added a color specification for "Flag Description" variable, which describes how the value was calculated/estimated. The scatterplot reveals that the production value of eggs increased from 1960 to 2020, with four countries experiencing much greater growth in production value than the others. In addition, the color specification reveals that the majority of values were calculated from unofficial data, making it difficult to discern if this variable had an effect.

```{r}

ggplot(egg_production, aes(x=Year, y=Value, color=`Flag Description`)) +
  geom_point(size=.5)+
  labs(title = "Production of Eggs by Year", x = "Year", y = "Production (tonnes)")

```
## Visualizing Part-Whole Relationships over Time

I chose to use a scatterplot with the production values of areas with values above 5,000,000 in order to compare the values of the largest producing areas. I used the color function to specify year in order to see the production value changed over time. This plot reveals that Asian regions, including "Asia", "China, mainland", and "Eastern Asia" experienced the greatest growth in egg production from 1960 to 2020.

```{r}
egg_new <- na.omit(egg_new)
egg_new <- egg_new[egg_new$Element == "Production",]
egg_large <- egg_new[egg_new$Value > 5000000,]
ggplot(egg_large, aes(x=Area,y=Value, color = Year)) + 
    geom_point() +
    theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))
    labs(title = "Production of Eggs by Year", x = "Area", y = "Production (tonnes)")
  
```

