---
title: "Challenge 5"
author: "Daniel Manning"
description: "Introduction to Visualization"
date: "1/9/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_5
---

```{r}
#| label: setup
#| warning: false
#| message: false
library(tidyverse)
library(ggplot2)
library(here)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


## Read in data

```{r}
cereal <- here("posts","_data","cereal.csv")%>%
  read_csv()
cereal
```

## Univariate Visualization 

I chose to use two histograms to display the Sodium and Sugar content of various cereal brands because these are single numeric variables.

```{r}
ggplot(cereal, aes(Sodium)) + 
  geom_histogram(aes(y = ..density..), alpha = 0.5) +
  geom_density(alpha = 0.2, fill="red")
  labs(title = "Sodium Content by Cereal Brand", x = "Sodium")

ggplot(cereal, aes(Sugar)) + 
  geom_histogram(aes(y = ..density..), alpha = 0.5) +
  geom_density(alpha = 0.2, fill="red")
  labs(title = "Sugar Content by Cereal Brand", x = "Sugar")
```
## Bivariate Visualization

I chose to use a scatterplot to display the Sodium by Sugar content of various cereal brands because these are numeric variables and the scatterplot can be used to reveal relationships between them.

```{r}
ggplot(cereal, aes(Sugar, Sodium)) + geom_point()
  labs(title = "Sodium Content by Sugar for Cereal Brands", x = "Sugar", y = "Sodium")
```

