---
title: "Challenge 4"
author: "Priya Marla"
description: "More data wrangling: mutate"
date: "01/12/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_4
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(lubridate)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  identify variables that need to be mutated
4)  mutate variables and sanity check all mutations

## Read in data

Read in one (or more) of the following datasets, using the correct R package and command.

-   abc_poll.csv ⭐
-   poultry_tidy.csv⭐⭐
-   FedFundsRate.csv⭐⭐⭐
-   hotel_bookings.csv⭐⭐⭐⭐
-   debt_in_trillions ⭐⭐⭐⭐⭐

```{r}
dataset <- read_csv("_data/hotel_bookings.csv") #reading the data from csv file
dataset
dim(dataset) #to get the number of rows and columns
unique(dataset$hotel) #to get the unique rows in hotel column
```

### Briefly describe the data
This dataset describes the information of the reservations made in Resort Hotel and City Hotel. There are 119390 rows and 32 columns. Each columns represents various data such as for what dates and what hotel the booking was made, whether reservation is called or not and if the payment is made or not etc.


## Tidy Data (as needed)

Is your data already tidy, or is there work to be done? Be sure to anticipate your end result to provide a sanity check, and document your work here.

The data needs to be cleaned. The date is spread across in multiple columns. If is there in a single column it;'ll will be easy to calculate stats. Can also calculate the date with lead time and date of arrival columns.


```{r}
select(dataset, 3:7 )
```

Any additional comments?

## Identify variables that need to be mutated

Are there any variables that require mutation to be usable in your analysis stream? For example, are all time variables correctly coded as dates? Are all string variables reduced and cleaned to sensible categories? Do you need to turn any variables into factors and reorder for ease of graphics and visualization?

Document your work here.

The country column is placed in the 14th place, it should be moved up to 2nd place i.e after the hotel column. Created a new column arrival_date by getting the date from "arrival_date_day_of_month", "arrival_date_month","arrival_date_year" columns. Removing these 3 columns and moving the arrival date column to column after "lead_time". 
Creating the new column booking_date to know the date of booking with the information from columns lead_time and arrival_date.
After tidying up the data, total columns remaining are 29

```{r}
#tidying the dataset
tidy_data <- dataset %>%
  relocate("country",.after = "hotel") %>% #relocating the country column
  mutate(arrival_date = (str_c(arrival_date_day_of_month,arrival_date_month,arrival_date_year, sep = "/")), arrival_date = dmy(arrival_date), .after = lead_time) %>% #variable for arrival date
  mutate(booking_date = arrival_date-days(lead_time), .after = lead_time) %>% #variable to know the date of booking
  select(-"lead_time")
tidy_data <- tidy_data[,-6:-9] #removed columns with arrival date information
tidy_data
```

Any additional comments?
