---
title: "Challenge 4"
author: "Henry Mitrano"
description: "More data wrangling: mutate"
date: "1/20/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_4
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(lubridate)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  identify variables that need to be mutated
4)  mutate variables and sanity check all mutations

## Read in data

```{r}
hotel_bookings = read_csv("_data/hotel_bookings.csv")
head(hotel_bookings)
```

### Briefly describe the data
It appears this is some kind of data regarding hotels in different countries, and statistics about their bookings. These statistics include number of people in a reservation, whether they're kids or adults, how their room was booked, whether they've been there before, etc. The data isn't organized horribly, but some stuff definitely stands out as needing tidying.


```{r}
tidy_bookings <- 
  hotel_bookings %>%
      mutate(arrival=str_c(arrival_date_day_of_month, arrival_date_month,arrival_date_year, sep="/"), arrival=dmy(arrival)) 

tidy_bookings = select(tidy_bookings, select = -starts_with("arrival_date") )



```

Any additional comments?

## Identify variables that need to be mutated

For starters, one of the more confusing things I found when looking through the data was that the arrival dates of the visits are kept in 3 separate columns (arrival date year, arrival date month, and arrival date day), while the checkout or "reservation status" date is stored in 1 column, normally. I'll want to replicate that mm-dd-yyyy format in one column for arrivals. We can doing that by importing an R library at the top of the file, "lubridate", that includes a bunch of date formatting tools.

```{r}
tidy_bookings<-transform(tidy_bookings,valued_customer=ifelse(is_repeated_guest >0 && previous_cancellations == 0, "yes", "no"))
head(tidy_bookings)
```
Additionally, I've added a column at the end that distingushes certain booking as important guests, based on historical data. If a customer has been booked at the hotel before and is a repeat visitor, and has never canceled a booking, they are listed as "yes" under the "valued_customer" column at the end of the data. 