---
title: "Challenge 8"
author: "Yoshita Varma Annam"
description: "Joining Data"
date: "1/16/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_8
  - railroads
  - snl
  - faostat
  - debt
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in multiple data sets, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  mutate variables as needed (including sanity checks)
4)  join two or more data sets and analyze some aspect of the joined data

(be sure to only include the category tags for the data you use!)

## Read in data

Read in one (or more) of the following datasets, using the correct R package and command.

  - military marriages ⭐⭐
  - faostat ⭐⭐
  - railroads  ⭐⭐⭐
  - fed_rate ⭐⭐⭐
  - debt ⭐⭐⭐
  - us_hh ⭐⭐⭐⭐
  - snl ⭐⭐⭐⭐⭐

```{r}
country_faostatcsv <- read_csv("_data/FAOSTAT_country_groups.csv")
cattle_faostatcsv <- read_csv("_data/FAOSTAT_cattle_dairy.csv")
```

### Briefly describe the data

```{r}
country_faostatcsv
summary(country_faostatcsv)
cattle_faostatcsv
summary(cattle_faostatcsv)
```

I have chosen country faostat and cattle faostat data set. After reviewing the data set it explains that country faostat dataset is codebook for country. This dataset has all the information country level. After joining this dataset with cattle faostat dataset we can perform data analysis based on different groups. 

In cattle dataset there is a lot of information about agriculture across the world. As you can see from the summary in element code there is information about milk, yield and production. Also, there is a common column in both the datasets country code and area code. We will be using this to join the data.


## Tidy Data (as needed)


```{r}
cattle_new <- rename (cattle_faostatcsv, "Country Code"= "Area Code" )

cattle_new
```

## Join Data

As the data is ready to join, we will use left_join to join the dataset using country code.

```{r}
final_data <- left_join(cattle_new, country_faostatcsv, by = "Country Code" )
final_data
```

```{r}
final_data_new <- rename (final_data, "Country_Group"= "Country Group" )
```


```{r}
final_data_new3 <- final_data_new%>%
  filter(Year >= 1998) %>%
  filter(`Country_Group`=="Central Asia") %>%
  filter(`Unit` == "tonnes") 
```

```{r}
final_data_new3 %>%
  ggplot(aes(x = Area, y = Value/100000, fill = Area)) +
  geom_bar(stat = "identity") + 
  labs(x = "Countries", y = "Different Agricultural Aspects", title = "Agricultural information across the world",
  fill = "Element")
```

```{r}
ggplot(final_data_new3) +
  geom_bar(aes(x = Area, fill= Element )) + 
  labs(x = "Countries", y = "Different Agricultural Aspects", title = "Agricultural information across the world",
  fill = "Element")

```




