---
title: "Challenge 9 Submission"
author: "Tanmay Agrawal"
description: "Creating a function"
date: "1/31/2023"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
    df-print: paged
categories:
  - challenge_9
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is simple. Create a function, and use it to perform a data analysis / cleaning / visualization task:

Examples of such functions are:
1) A function that reads in and cleans a dataset.  
2) A function that computes summary statistics (e.g., computes the z score for a variable).  
3) A function that plots a histogram.  

That's it!

## Reusable Functions!

My goal here to create generic function templates that work with majority of the datasets. These functions can then be reused and modified based on specific requirements of the dataset. 

```{r}
# A function that reads in and cleans a dataset
read_and_clean_data <- function(file) {
  data <- read.csv(file)
  
  # Remove missing values
  data <- na.omit(data)
  return(data)
}

# A function that computes summary statistics
compute_summary_stats <- function(data, variable) {
  
  # Get statistics for only numeric types
  numeric_data <- select_if(data, is.numeric)
  general_stats <- summary(numeric_data)
  
  # Compute the mean and standard deviation separately for the Z-Score
  mean_value <- mean(data[, variable])
  std_value <- sd(data[, variable])
  
  # Compute the z score for the variable
  z_score <- (data[, variable] - mean_value) / std_value
  
  # Return the z score
  return(list(general_stats, z_score))

}

# A function that plots a histogram

plot_histogram <- function(data, variable) {
  
  hist(as.numeric(data[, variable]), main = paste("Histogram of", variable), xlab = variable, col = "blue", breaks = "Sturges")
}



```

Now that we have our templates ready, let's try it on our dataset, the old faithful: `hotel_bookings`.

```{r}
# Read and clean the data
data <- read_and_clean_data("_data/hotel_bookings.csv")
# As opposed to: data = read_csv("_data/hotel_bookings.csv")

# Compute the z score for a variable
summary_stats <- compute_summary_stats(data, "stays_in_week_nights")

summary_stats[1] # General stats
z_scores <- summary_stats[2]

# Plot the histogram of the variable
plot_histogram(data, "stays_in_week_nights")

```

