---
title: "Challenge8_MarcelaRobinson"
author: "Marcela Robinson"
editor: visual
Date: January 22, 2023
Description: "Reading multiple data sets"
Categories:
 - Challenge_8
 - FAOSTAT 
 - cattle
 - egg_chicken

---

```{r}
library(here)
library(tidyverse)
```

```{r}
#Read the first dataset
cattle<- here("posts", "_data", "FAOSTAT_cattle_dairy.csv")%>%
read_csv()

#Read second dataset 
chicken<-here("posts", "_data", "FAOSTAT_egg_chicken.csv")%>%
read_csv()
```

For this challenge, I'll be joining two sets of data. Both sets of data are provided by the Food and Agriculture Organization of the United Nations. The first set, **cattle**, represent the vaues of livestock (Milk Animals, Yield, or Production) in different areas of the world over the years. Other variables include unit, flag, flag description, and different codes. 

The second set of data that I'll be joining, **chicken**, contains values of chickens (Laying, Yield, or Production) in different areas of the world over the years. 

```{r}

#Join both datasets 
joined<-bind_rows(cattle, chicken)
print(joined)

```
I'm using the bind_rows function since both datasets have the same variables. As I result, my new dataset called **joined** has 74619 observations and 14 variables. 

## Clean the data

```{r}
#Remove unnecessary columns 
joined%>%
  select(-c("Domain Code", "Domain", "Area Code","Element Code", "Flag", "Flag Description", "Year Code", "Item Code"))

print(joined)
```


