---
title: "Challenge 6"
author: "Priya Marla"
description: "Visualizing Time and Relationships"
date: "1/19/2023"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_6
  - hotel_bookings
  - air_bnb
  - fed_rate
  - debt
  - usa_hh
  - abc_poll
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(readxl)
library(ggplot2)
library(tseries)
library(here)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  mutate variables as needed (including sanity checks)
4)  create at least one graph including time (evolution)
   - try to make them "publication" ready (optional)
   - Explain why you choose the specific graph type
5)  Create at least one graph depicting part-whole or flow relationships
   - try to make them "publication" ready (optional)
   - Explain why you choose the specific graph type

[R Graph Gallery](https://r-graph-gallery.com/) is a good starting point for thinking about what information is conveyed in standard graph types, and includes example R code.

(be sure to only include the category tags for the data you use!)

## Read in data

Read in one (or more) of the following datasets, using the correct R package and command.

  - debt  ⭐
  - fed_rate ⭐⭐
  - abc_poll ⭐⭐⭐
  - usa_hh ⭐⭐⭐
  - hotel_bookings ⭐⭐⭐⭐
  - air_bnb  ⭐⭐⭐⭐⭐

```{r}
dataset <- here("posts","_data","USA Households by Total Money Income, Race, and Hispanic Origin of Householder 1967 to 2019.xlsx") %>%
read_excel(skip=5, n_max = 351,
           col_names = c("Year", "Population", "del","Income_1", "Income_2", "Income_3", "Income_4","Income_5","Income_6","Income_7","Income_8","Income_9","Median_Income", "Median_error","Mean_Income", "Mean_error"))
dataset
```

### Briefly describe the data
This dataset describes the income statistics of various races over the years. The data shows what percentage of people fall under each income category i.e ranging from under $15,000 to $200,000 per year. The mean and median of the income is also calculated in the columns 13 and 15. The statistics are calculated all races combine and also separate for different races.


## Tidy Data (as needed)

Is your data already tidy, or is there work to be done? Be sure to anticipate your end result to provide a sanity check, and document your work here.

There are few unnecessary columns in the data like Total, which has to be deleted. A new column which represents the race has to mutated to the data. Superscripts have to be removed from the year column.

```{r}

household <- dataset %>%
  mutate(Race = case_when(is.na(Population) ~ Year), .before = Year) %>% #Created a new column race
  fill(Race, .direction = "down") %>% #filled the value in the race column
  separate(Year, into=c("Year", "del"), sep=" ")%>% #Removing the extra characters in the Year column
  select(!contains("del")) %>% #deleting unecessary columns
  drop_na() #dropping null values if any

household$Race <- gsub('[0-9]+', '', household$Race) #Removing extra characters in the Race column

household
```

Are there any variables that require mutation to be usable in your analysis stream? For example, do you need to calculate new values in order to graph them? Can string values be represented numerically? Do you need to turn any variables into factors and reorder for ease of graphics and visualization?

Added a new column race while tidying up the data, which is useful for facet_wrap function to plot different graphs on basis of the values in the Race column.
Converting the all the columns to numeric for ease during plotting the graphs. Converted the Year column to date format

Document your work here.
```{r}
#Converting all the columns except Race and Year to numeric for better 
for (i in names(household)){
  if (i == "Race"){
    next
  }
  else if (i == "Year"){
    next
  }
  else{
    household[i] <- sapply(household[i],as.numeric)
  }
}

#converting year to date format
household$Year <- as.Date(as.character(household$Year), format = "%Y") 

class(household$Year)
class(household$Mean_Income)
```


## Time Dependent Visualization

By plotting the Mean_income of various races over the time (years), we can see that there is an increasing trend of the mean over the years and so is the Median. 
The Mean income ranged between $30,000 to $120,000 whereas median ranged between $50,000 to $100,000.

```{r}
#Mean Income
ggplot(data = household, aes(Year, Mean_Income)) +
  geom_line(color="violet") +
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(title = "Plots for Mean Income of different race over the years") +
  facet_wrap(~ Race) +
  theme(
  strip.text = element_text(size = 6),
  panel.grid.minor = element_blank(),               #removes minor grid lines
  panel.grid.major = element_blank())
#Median Income
ggplot(data = household, aes(Year, Median_Income)) +
  geom_line(color="red") +
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(title = "Plots for Median Income of different race over the years") +
  facet_wrap(~ Race) +
  theme(
  strip.text = element_text(size = 6),
  panel.grid.minor = element_blank(),               #removes minor grid lines
  panel.grid.major = element_blank())
```

## Visualizing Part-Whole Relationships

In the below comparison graph of mean and median income of various races over the years, we ca refer that the asian community has logged the highest mean and median income. Least income is registered by the black community. 

```{r}
#Mean_income Plot
ggplot(data = household, aes(x = Year, y = Mean_Income, col = Race)) +
  geom_line()+
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(title = "Comparison of Mean Income of all races over the year")
#Median_income Plot
ggplot(data = household, aes(x = Year, y = Median_Income, col = Race)) +
  geom_line()+
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(title = "Comparison of Median Income of all races over the year")
```
