---
title: "Challenge 9 Yoshita Varma"
author: "Yoshita Varma"
description: "Creating a function"
date: "1/27/2023"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_9
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is simple. Create a function, and use it to perform a data analysis / cleaning / visualization task:

Examples of such functions are:
1) A function that reads in and cleans a dataset.  
2) A function that computes summary statistics (e.g., computes the z score for a variable).  
3) A function that plots a histogram.  

That's it!

## A function that reads in and cleans a dataset. 

```{r}
func_read_clean <- function(x) {
  data <- read_csv(x)
  data <- data %>% select_if(~ !any(is.na(.)))
  return(data)
} 
```

Reading hotel booking data as I am familiar with it. 

```{r}
hotel_booking <- func_read_clean("_data/hotel_bookings.csv")
hotel_booking
```


## A function that computes summary statistics (e.g., computes the z score for a variable).

```{r}
statistics <- function(x){
  stat <- tibble(
    mean=mean(x),
    median=median(x),
    sd=sd(x)
  )
  return(stat)
}
statistics(hotel_booking$stays_in_week_nights)
```
Printing stats of stays_in_week_nights.

## A function that plots a histogram.

```{r}
histogram <- function(data_read){
  HotelBookings_csv_mutate <- data_read %>% 
  mutate(arrival_date = str_c(arrival_date_year, 
                              arrival_date_month, arrival_date_day_of_month, sep="/"),
         arrival_date = lubridate::ymd(arrival_date)) %>% 
  select(-c(arrival_date_year,arrival_date_month, arrival_date_day_of_month))

  
ggplot(HotelBookings_csv_mutate, aes(x=arrival_date, y= stays_in_week_nights, color = `hotel`)) + 
  geom_line() + 
  xlab("Year") + 
  ylab("Number of days stay during week") + 
  ggtitle("Year vs Stays_in_week_nights")
}

histogram(hotel_booking)
```

This shows some might book a hotel instead of renting a place. One common thing observed is the spike in the start of the year.


