---
title: "Challenge 5"
author: "Paarth Tandon"
description: "Introduction to Visualization"
date: "1/9/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
    df-print: kable
categories:
  - challenge_5
  - cereal
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Read in data

Read in one (or more) of the following datasets, using the correct R package and command.

-   cereal ⭐
-   pathogen cost ⭐
-   Australian Marriage ⭐⭐ 
-   AB_NYC_2019.csv ⭐⭐⭐
-   railroads ⭐⭐⭐
-   Public School Characteristics ⭐⭐⭐⭐ 
-   USA Households ⭐⭐⭐⭐⭐

```{r}
set.seed(42)
# read in the data using readr
cereal <- read_csv("_data/cereal.csv")
head(cereal, 5)
```

### Briefly describe the data

This data set contains four columns:

* `Cereal <chr>`: The name of the cereal
* `Sodium <dbl>`: The amount of sodium in a serving of the cereal
* `Sugar <dbl>`: The amount of sugar in a serving of cereal
* `Type <chr>`: The type of the cereal (Child or Adult)

## Tidy Data (as needed)

Data is already tidy. Going to mutate `Type` for visualization reasons.

```{r}
cereal <- mutate(cereal, `Type` = recode(`Type`,
    "A" = "Adult",
    "C" = "Child"
))
head(cereal, 5)
```

## Univariate Visualizations

```{r}
ggplot(cereal, aes(`Sodium`)) +
geom_histogram(binwidth = 100) +
facet_grid(vars(`Type`)) +
labs(title = "Distribution of Sodium based on Type")
```

```{r}
ggplot(cereal, aes(`Sugar`)) +
geom_histogram(binwidth = 5) +
facet_grid(vars(`Type`)) +
labs(title = "Distribution of Sugar based on Type")
```

I chose histograms, since we are working with univariate, continuous data. I wanted to compare by type, so I applied a facet_grid.

## Bivariate Visualization(s)

```{r}
ggplot(cereal, aes(`Type`, `Sugar`)) +
geom_boxplot() +
labs(title = "Amount of Sugar in Cereal by Type")
```

Here we can see that children's cereal has more sugar in it than adult's cereal, on average. I chose a box plot, since the data is very sparse. With only 20 points, we can only extrapolate so much.
