---
title: "Challenge 4"
author: "Yoshita Varma Annam"
description: "More data wrangling: mutate"
date: "1/2/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_4
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  identify variables that need to be mutated
4)  mutate variables and sanity check all mutations

## Read in data

Read in one (or more) of the following datasets, using the correct R package and command.

-   abc_poll.csv ⭐
-   poultry_tidy.csv⭐⭐
-   FedFundsRate.csv⭐⭐⭐
-   hotel_bookings.csv⭐⭐⭐⭐
-   debt_in_trillions ⭐⭐⭐⭐⭐

```{r}
library(readr)

HotelBookings_csv <- read_csv("_data/hotel_bookings.csv")
```
As I am already familiar with hotel booking data from challenge 2, I choose this dataset to work on in challenge 4. Because in the challenge 2 I some part of the data wasn't very clean and can be mutated for better analysis. 

### Briefly describe the data

Some of my analysis is based on challenge 2. Reflecting on those please refer below.
```{r}
#| label: understanding the data
HotelBookings_csv
```

By just viewing the data it looks like the data is about 119,390 hotel entries and detailing for 32 features. The features mainly describe the booking entirely based on their arrival, cancellations and timings. It also accounts the number of babies, children, adults across the world. There is a separate field to verify for the repeated guests.  To understand further we need to perform more operations.

```{r}
#| label: summary of the data
summary(HotelBookings_csv)
```
```{r}
#| label: column names the data
colnames(HotelBookings_csv)
```
```{r}
#| label: finding unique values of the data
unique(HotelBookings_csv$deposit_type)
length(unique(HotelBookings_csv$market_segment))
unique(HotelBookings_csv$market_segment)
length(unique(HotelBookings_csv$market_segment))
unique(HotelBookings_csv$distribution_channel)
length(unique(HotelBookings_csv$distribution_channel))
unique(HotelBookings_csv$hotel)
length(unique(HotelBookings_csv$hotel))
unique(HotelBookings_csv$country)
length(unique(HotelBookings_csv$country))
```

After the following analysis it is clear that the data has been collected across the world for different countries approximately 150-180 from 2015 to 2017. The data is very specific to two kinds of hotels- "Resort Hotel", "City Hotel". There are majorly 8 kinds of bookings which include all the professional to personal types like- Corporate, Aviation etc. If we observe the mean from the summaries it can be said that there were approximately 185% adults, children 10%, and 1% baby have come to stay in the hotels. Similarly, on an average people stayed for 2.5 days during the week and 1 day during the weekends. The stats are only based on the summaries. To further colcude more accurately for this data need more analysis.

## Tidy Data (as needed)


```{r}
#Null values in country column
table(HotelBookings_csv$country)

```

After removing the rows which has NULL values in country coulmn as this is irrelevant to our analysis.

```{r}
# Removing Null values
HotelBookings_csv <- HotelBookings_csv %>% 
  filter(!(country == "NULL"))
```
 
As arrival year, month, date has been stored in three different columns we can combine the arrival date in year, month and the date into one single field and name it as the arrival date. Also, I feel arrival_date_week_number is irrelevant for my analysis as we can take a good guess of the week number from the new field arrival date.

```{r}
# Remove Columns by Index
HotelBookings_csv <- HotelBookings_csv[,-6]
HotelBookings_csv
```
Any additional comments?

## Identify variables that need to be mutated

Are there any variables that require mutation to be usable in your analysis stream? For example, are all time variables correctly coded as dates? Are all string variables reduced and cleaned to sensible categories? Do you need to turn any variables into factors and reorder for ease of graphics and visualization?

As the data has been modified as per required I'll go ahead and mutate it to have single field for arrivale date as mentioned above.

```{r}
#Mutating the arrival date into a single field 

HotelBookings_csv_mutate <- HotelBookings_csv %>% 
  mutate(arrival_date = str_c(arrival_date_day_of_month,
                              arrival_date_month,
                              arrival_date_year, sep="/"),
         arrival_date = lubridate::dmy(arrival_date)) %>% 
  select(-c(arrival_date_day_of_month,arrival_date_month,arrival_date_year))

HotelBookings_csv_mutate
```
Similarly,  changing the datatype of the columns of company and agent as they have numerical values due to some NULL values it has character as datatype. First handling NULL values and changing them NA as the targetted datatype will be numeric.

```{r}
#Mutating the class of the agent and company field from character to numeric

HotelBookings_csv_mutate <- HotelBookings_csv_mutate %>%
  mutate(across(c(agent, company),~ replace(.,str_detect(., "NULL"), NA))) %>% mutate_at(vars(agent, company),as.numeric)

```

Now reviewing the summary of the mutated data. 

```{r}
summary(HotelBookings_csv_mutate)
```

As you can see there is only one field for arrival date and dataype for columns agent and company has been changed to Numeric.
