---
title: "HW 2"
author: "Daniel Manning"
description: "Reading in Data"
date: "1/7/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(here)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Tidying of Dataset

First I loaded the "FAOSTAT_cattle_dairy.csv dataset. This dataset was already largely tidy, but I finished the process by removing variables that were redundant or only had one case, leaving each column with its own variable and each row as an occurrence. 


```{r}
cattle_dairy <- here("posts","_data","FAOSTAT_cattle_dairy.csv")%>%
  read_csv()
cattle_dairy
cattle_dairy_new <- cattle_dairy %>% 
  select(-c('Domain', 'Domain Code', 'Area Code', 'Element', 'Element Code', 'Item Code', 'Year Code', 'Unit', 'Flag'))
cattle_dairy_new
```


## Narrative, Variables, and Research Question

This dataset consists of five variables with the following types:
Area: string
Item: string
Year: double
Value: double
Flag Description: string

The overall dataset represents values for whole fresh cow milk in different countries across different years, as well as whether the data was estiamted/calculated. 

One potential research question could investigate the difference in values with respect to the area. For example: Which area produced the most whole fresh cow milk? What country produced the least whole fresh cow milk? What was the distribution of average whole fresh cow milk production for each area?

Another research question could investigate the change in milk production with respect to year. For example: Did the production of whole fresh cow milk increase from 1961 to 2018? During which year was the most cow milk produced and during which year was the least produced? 

A third research question could investigate the relationship between the method of estimating/calculating values and the values themselves. For example: Is there a difference between the average of values described as "Calculated data" and "FAO estimates". 

Lastly, this dataset could be used to investigate relationships between multiple variables, such as year and area. For example: Did some areas see an increase in production from 1961 to 2018 while others saw a decrease? Was the magnitude of change in production different across areas? 
