---
title: "MarcelaRobinson_challenge4"
author: "Marcela Robinson"
editor: visual
df-print: paged
---

```{r}
library(tidyverse)
library(lubridate)
library(stringr)

```

##Read in data 
```{r}
poultry<-read_csv("_data/poultry_tidy.csv")
head(poultry)
```
The dataset poultry contains information related to the price per month of poultry cuts from 2004 - 2013. The dataset contains 600 observations and 4 different variables. The variables for this dataset are:
-Product: This variable contains different poultry cuts including whole, breast, boneless breast, bone-in breast, whole legs, and thighs. 
-Year and month: These variables contain the month and the year for each poultry cut price. 
Price_dollar: The dollar amount per month for each poultry cut.

The first step I want to do to clean the data is to combine the month and the year to reduce the number of columns. Next, I would like to use the mutate_if function to have 2 decimals for the price of the cuts.

##Combine year and month to reduce number of columns
```{r}
poultry_date<-poultry%>%
  mutate(date = str_c(Month, Year, sep=" "),
         date = my(date))%>%
  select(Product, Price_Dollar,date)

head(poultry_date)

```

##Move decimal place 
```{r}
poultry_date%>%
  mutate_if(is.numeric, ~round(., 2))
```


