---
title: "Challenge 3"
author: "Siddharth Goel"
description: "Tidy Data: Pivoting"
date: "01/27/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_3
  - animal_weights.csv
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Read in data
```{r}
animal_weights_df <- read_csv("_data/animal_weight.csv")

animal_weights_df
spec(animal_weights_df)
```


### Briefly describe the data

The input dataset describes the weights of different animals and the overall category-wise weights for 9 IPCC areas. For this dataset, I plan to move the animals to rows and the areas to columns using pivot_longer so that it is easy to get statistics based on specific animals and categories. 

### Challenge: Describe the final dimensions

Document your work here.

```{r}
# Existing Rows
expected_rows <- nrow(animal_weights_df) # 9

# Existing Columns
expected_cols <- ncol(animal_weights_df) # 17

# Expected Rows
nrow(animal_weights_df) * (expected_cols - 1) # -1 for the header

# expected columns 
3 # IPCC Area, Animal, Weight
```

The dataset currently has 9 rows and 17 columns.
When we pivot the dataset, the targeted rows and columns will be 144 rows and 3 columns (Area, Animal and Weight). 

## Pivot the Data

Now we will pivot the data, and compare our pivoted data dimensions to the dimensions calculated above as a "sanity" check.

```{r}
pivoted_df <- pivot_longer(animal_weights_df, 2:ncol(animal_weights_df), names_to = "Animal", values_to = "Weight")
head(pivoted_df)
```
Yes, once it is pivoted long, our resulting data are $144 x 3$ - exactly what we expected!
