---
title: "Challenge 1"
author: "Siddharth Goel"
description: "Reading in data and creating a post"
date: "01/20/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(readr)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

```{r}
birds_set = read_csv('_data/birds.csv')
spec(birds_set)
head(birds_set)

```

This dataset has 14 columns and 30977 data values. All the columns are either of the type `col_character` or `col_double`


## Describe the data
From the columns and the data, we can see that there are multiple columns that represent the same data in multiple forms. For example, `Area` and `Area code`, `Domain` and `Domain Code`, `Element` and `Element Code`, `Item` and `Item Code`, and `Year` and `Year Code`. We can de-duplicate these columns and create separate mappings to reduce the size of the data.
Also, the columns `Domain` and `Element` have a single value which means that these columns can also be eliminated.

```{r}
#| label: summary
unique(birds_set$Domain)
length(unique(birds_set$Area))
unique(birds_set$Item)
unique(birds_set$Element)
length(unique(birds_set$Year))
unique(birds_set$Flag)
```

By analyzing the data values and the unique column values, we can assert that the dataset contains the livestock data about five birds over a certain period of time. This dataset is over a single domain, which is `Live Animals` and mainly contains information about birds and the regions they belong to. 