---
title: "Challenge 5"
author: "Siddharth Goel"
description: "Introduction to Visualization"
date: "01/29/2023"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_5
  - air_bnb
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(readr)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  mutate variables as needed (including sanity checks)
4)  create at least two univariate visualizations
   - try to make them "publication" ready
   - Explain why you choose the specific graph type
5)  Create at least one bivariate visualization
   - try to make them "publication" ready
   - Explain why you choose the specific graph type


## Read in data

```{r}
data <- read.csv("_data/AB_NYC_2019.csv")
head(data)
```

### Briefly describe the data
This dataset contains data for AirBNB hotels and rooms available for booking in New York City for the year 2019. It has 16 columns with details ranging from booking dates, locations to room type and reviews.

## Tidy Data (as needed)

Is your data already tidy, or is there work to be done? Be sure to anticipate your end result to provide a sanity check, and document your work here.

I am tidying the data to remove the N/A values in `reviews_per_month` column with 0.

```{r}
tidy_data <- data %>% replace_na(list(reviews_per_month = 0))
```

Are there any variables that require mutation to be usable in your analysis stream? For example, do you need to calculate new values in order to graph them? Can string values be represented numerically? Do you need to turn any variables into factors and reorder for ease of graphics and visualization?

```{r}
head(tidy_data)
```
All the variables are in correct format and are usable in the analysis.

## Univariate Visualizations
I will attempt to plot the reviews_per_month of listings in NYC as this is the closest metric we can get for occupancy in these listings considering occupancy to be directly proportional to reviews.

To achieve this plot, I first aggregated the data to get the total reviews in a month for each neighborhood group.

```{r}
df_agg = tidy_data %>% group_by(neighbourhood_group)  %>%
                    summarise(total_reviews_per_month = sum(reviews_per_month),
                              .groups = 'drop')
head(df_agg)
ggplot(data=df_agg, aes(x=neighbourhood_group, y=total_reviews_per_month)) +
  geom_bar(stat="identity") + labs(title = "Total Reviews Per Month")

```


## Bivariate Visualization(s)
The pricing for each borough is also an important factor that is generally looked upon by customers before booking and is worth keeping track of.

```{r}
tidy_data %>%
  ggplot(aes(neighbourhood_group, price), fill = neighbourhood) + 
  geom_boxplot() + 
  labs(title = "Borough-wise pricing") + 
  theme_classic()
```