---
title: "Challenge 9"
author: "Siddharth Goel"
description: "Creating a function"
date: "01/31/2023"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_9
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(dplyr)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is simple. Create a function, and use it to perform a data analysis / cleaning / visualization task:

Examples of such functions are:
1) A function that reads in and cleans a dataset.  
2) A function that computes summary statistics (e.g., computes the z score for a variable).  
3) A function that plots a histogram.  

That's it!

# Function to read and clean a dataset
This function reads in a csv and removes the N/A columns in the file

```{r}
read_and_clean_csv <- function(x) {
  data <- read_csv(x)
  data <- data %>% select_if(~ !any(is.na(.)))
  return(data)
} 
```

Running the function on the dataset hotel_bookings.csv
```{r}
data <- read_and_clean_csv("_data/hotel_bookings.csv")
head(data)
```

# Function to compute statistics (z-score)
z-score = (x - mean) / std_dev

```{r}
z_score <- function(col) {
  output <- (col - mean(col)) / sd(col)
  return(output)
}
```


Evaluating Z-Score for a column in hotel_bookings.csv
```{r}
z_score_output <- z_score(data$stays_in_week_nights)
head(z_score_output)
```


# Function that plots a histogram

```{r}
plot_histogram <- function(input_path) {
  tidy_data <- read_csv(input_path)
  tidy_data <- tidy_data %>% replace_na(list(reviews_per_month = 0))
  df_agg = tidy_data %>% group_by(neighbourhood_group)  %>%
                    summarise(total_reviews_per_month = sum(reviews_per_month),
                              .groups = 'drop')
  head(df_agg)
  ggplot(data=df_agg, aes(x=neighbourhood_group, y=total_reviews_per_month)) +
  geom_bar(stat="identity") + labs(title = "Total Reviews Per Month")
}

plot_histogram("_data/AB_NYC_2019.csv")
```






