---
title: "Challenge 5: Introduction to Visualization"
author: "Surya Praneeth Reddy Chirasani"
description: ""
date: "01/15/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_5
  - cereal
---

```{r}
#| label: setup
#| warning: false
library(tidyverse)
library(readr)
knitr::opts_chunk$set(echo = TRUE)
```

## Data Description
The dataset includes sodium and sugar content for 20 cereals which are categorized into two types A or C. There isn't any information regarding what the types mean

```{r}
cereal_data <-read_csv("_data/cereal.csv")
cereal_data
```
## Tidying Data
Looking at the entire data I don't see the need for tidying data

## Univariate Visualization

I want to plot a histogram to see in what ranges or distribution sodium and sugar contents lie

```{r}
ggplot(cereal_data, aes(x=Sodium)) +
  geom_histogram(bins=10) 
```

```{r}
ggplot(cereal_data, aes(x=Sugar)) +
  geom_histogram(bins=10) 
```

In the histogram for Sodium there is missing data in some bins, which could mean there are outliers. We can identify them by plotting a box plot. So for this reason now I am plotting a box plot

```{r}

is_outlier <- function(x) {
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}

cereal_data %>% 
  mutate(outlier = ifelse(is_outlier(Sodium), Cereal, as.numeric(NA))) %>% 
  mutate(Cereal= 'Cereal') %>%
  ggplot(., mapping=aes(x=Cereal, y=Sodium))+geom_boxplot()+ geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3)
```
Indeed there are outliers in the boxplot!

```{r}

is_outlier <- function(x) {
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}

cereal_data %>% 
  mutate(outlier = ifelse(is_outlier(Sugar), Cereal, as.numeric(NA))) %>% 
  mutate(Cereal= 'Cereal') %>%
  ggplot(., mapping=aes(x=Cereal, y=Sugar))+geom_boxplot()+ geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3)
```
The histogram for sugar didn't have any empty bins and so as observed in the boxplot there are no outliers

## Bivariate Visualizations

In order to observe the relationship between Sodium and Sugar, I want to plot a scatter plot with Sodium on y-axis and Sugar in x-axis and see if there are some cereals which has high sugar and low in sodium and viceversa

```{r}
ggplot(cereal_data, aes(Sugar , Sodium)) + geom_point()
```

Looking at the scatterplot, there are all kinds of cereals with no particular relationship. Maybe there could be a relationship when we highlight cereals based on their type 

```{r}
is_outlier <- function(x) {
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}

cereal_data %>% 
  mutate(outlier = ifelse(is_outlier(Sugar) | is_outlier(Sodium), Cereal, as.numeric(NA))) %>% 
  ggplot(., aes(x = Sugar, y = Sodium, color=Type)) + theme(plot.title = element_text(hjust = 0.5)) +
  geom_point() + 
  scale_colour_manual(values = c("blue", "red")) +
  geom_text(aes(label = outlier), color="black", na.rm = TRUE, hjust=1.1, vjust=1.1, size = 2) +
  labs(title = "Scatter Plot", x = "Sugar", y = "Sodium")
```
There is no particular relationship between cereals of each type as there are spread out in the graph. But plotting the scatter plot helped in visualizing the data more clearly

