---
title: "Challenge 9: Creating a Function"
author: "Surya Praneeth Reddy Chirasani"
description: ""
date: "01/29/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_9
  - debt
---

```{r}
#| label: setup
#| warning: false
library(tidyverse)
library(readr)
library("readxl")
library(lubridate)
knitr::opts_chunk$set(echo = TRUE)
```

## Function that reads a dataset and cleans it

Created a function to read in the "Debt in Trillions" dataset and create a new column with readable dates from the "Year and Quarter" Column

```{r}
debt_data <- function(x) {
  data <-read_excel("_data/debt_in_trillions.xlsx")
  data <- data%>%
        mutate(Date = parse_date_time(`Year and Quarter`, orders="yq"),
        .before=`Year and Quarter`)
  return(data)
} 
```

```{r}
debt_data()
```


## Function to calculate summary statistics

I have used the debt dataset for this task as it has lot of numeric data

```{r}
data <-read_excel("_data/debt_in_trillions.xlsx")
```

I created a function to calculate z score for a variable. It takes a vector as input, calculates z score for each value in the vector and returns the vector as output

```{r}
z_score <- function(x){
  stat <- (x - mean(x,na.rm=T))/sd(x,na.rm=T)
  return(stat)
}
z_score(data$Mortgage)
```
