---
title: "Challenge 3: Pivoting"
author: "Surya Praneeth Reddy Chirasani"
desription: "Pivoting data in a dataset"
date: "01/07/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge3
  - tidydata
  - pivot_longer
---

```{r}
#| label: setup
#| warning: false
library(tidyverse)
library(readr)
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
animal_weights <-read_csv("_data/animal_weight.csv", show_col_types = FALSE)
animal_weights
```
The data for each animal for each area is in one row, I want to pivot the data from the columns to rows using pivot_longer, so that I can group by each area and get summary statistics for each area. These statistics cannot be calculated easily in the present format, for this reason I want to pivot the data. There are 16 lifestock types across the 9 regions which when we pivot to rows, the total number of results will be 16*9=144 rows


```{r}
pivot_longer(animal_weights, 2:ncol(animal_weights), names_to = "animal", values_to = "weight")
```


After pivoting the data, the dimension of the tibble is 144x3 which is as expected

