---
title: "Challenge 8: Joining Data"
author: "Surya Praneeth Reddy Chirasani"
description: ""
date: "01/28/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_8
  - snl
---

```{r}
#| label: setup
#| warning: false
library(tidyverse)
library(readr)
knitr::opts_chunk$set(echo = TRUE)
```

## Data Description

For this challenge I am using SNL dataset for my analysis. There are three data files: actors.csv, casts.csv and season.csv. actors file contains gender information regarding different types("cast", "guest", "crew", "unknown") of actors, casts file contains information regarding which seasons the actor appeared in and how many episodes, and finally the seasons file has information about, what year the season of SNL happened and total number of episodes that season has.

```{r}
actors <- read_csv("_data/snl_actors.csv", show_col_types = FALSE)
casts <- read_csv("_data/snl_casts.csv", show_col_types = FALSE)
seasons <- read_csv("_data/snl_seasons.csv", show_col_types = FALSE)
```


```{r}
actors
```

```{r}
casts
```
```{r}
seasons
```

## Gender Analysis

Looking at the data, gender is one aspect we can see how the casts makeup changed with time. Since this is a known disparity in most of the industries including entertainment industry, this would be a good question to answer. For the analysis, instead of just comparing number of cast for each gender type in each season, to keep it fair I calculated the total number of episodes each gender appeared in each season. This metric seems better representation of how much each gender appeared on SNL

```{r}
casts_gender_episodes_count <- casts  %>%
  left_join(filter(actors,type=="cast"),
            by="aid") %>%
  count(sid,gender, n_episodes) %>%
  mutate(total_episodes = n_episodes*n) %>%
  select(sid, gender, total_episodes) %>%
  group_by(sid, gender) %>%
  summarise(total = sum(total_episodes))%>%
  ungroup() %>%
  left_join(seasons)%>%
  filter(!is.na(gender))%>%
  select(sid, gender, year, total)
casts_gender_episodes_count
```

I have used "left_join" to join the casts and actors data since we are focusing our analysis on gender makeup of "casts". After that, I counted the number of episodes each gender appeared in, using count function. I used "group_by" to sum over all the episodes all the actors of each gender appeared in. Finally, I have used "left_join" again to get year information of each season, which will help label the x-axis in visualization.

```{r}
casts_gender_prop <- casts_gender_episodes_count %>%
  group_by(sid) %>%
  mutate(prop=total/sum(total)) %>%
  ungroup() %>%
  select(-total)  %>%
  pivot_wider(names_from = gender,
              values_from = prop)%>%
  mutate_at(-1, ~replace_na(., 0))%>%
  pivot_longer(c(female,male,unknown),
               values_to = "prop",
               names_to =  "gender")
casts_gender_prop
```

In the next series of operations,I used "group_by" and "mutate" to compute the proportion of gender per season. Some seasons only have "male" and "female" as the unique genders, while others have "male", "female", "other", and “NA". To include all genders for each season, I used "pivot_wider", "mutate_at" to ensure each level of gender is available for each season. In the last step, I have used pivot_longer() to get the data back in a format that will be helpful for visualization.

## Visualization

```{r}
ggplot(casts_gender_prop,aes(year,prop,col=gender)) + theme(plot.title = element_text(hjust = 0.5)) +
  geom_line(linewidth=1)+
  labs(title = "Saturday Night Live - Gender Progression Analysis", x = "Year", y = "Proportion")
```
The gender makeup fluctuated a lot over the years slowly moving towards 0.5(equality) but far from it, still in 2020. With just the cast gender makeup and number of episodes they appeared in, the analysis I have presented is a simple one. A complex analysis will take into account how much time each gender spent in each episode doing multiple segments for example. Probably male gender could have done more segments compared to female gender and then the difference between the red and green lines will be even more than in the graph I have plotted above.

In terms of learning, this challenge helped me lot to learn join functions and what variables to consider in order to answer my research question