---
title: "Challenge 7: Visualizing Multiple Dimensions"
author: "Surya Praneeth Reddy Chirasani"
description: ""
date: "01/27/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_7
  - eggs
---

```{r}
#| label: setup
#| warning: false
library(tidyverse)
library("readxl")
library(ggplot2)
library(lubridate)
knitr::opts_chunk$set(echo = TRUE)
```

## Data Description

For this challenge, I have decided to work with eggs dataset. This dataset has monthly price amount of different sizes(large and extra large) of half dozen and dozen eggs. 

```{r}
eggs_data <-read_csv("_data/eggs_tidy.csv")
eggs_data
```
## Tidying Data

In order to plot a graph of varying prices for different sizes and amount of eggs, I need to pivot the data longer and separate the data based on size and amount. This is done as follows

```{r}
eggs_tidy <- eggs_data %>%
  pivot_longer(cols=3:6, names_to="type", values_to = "price") %>%
  mutate(type=str_replace(type,"extra_large","extra large"),
         type=str_replace(type,"half_dozen","half dozen")) %>%
  separate(type,into=c("size","amount"),sep="_") %>%
  mutate(date = str_c(year, month, sep=" ")) %>%
  mutate(date = ym(date))
eggs_tidy
```

```{r}
ggplot(eggs_tidy,aes(date,price)) +
  geom_line(color = "steelblue", size = 1)+
  scale_x_date()+
  scale_y_continuous(labels=scales::dollar_format(), limits=c(0,300)) +
  facet_grid(size ~ amount) +
  labs(x="Date",y="Average Price Per Carton")+
  theme(axis.text.x=element_text(angle=90))
```

Since we are just trying to visualize variation of price with time for different sizes and amount, I decided to plot a simple line graph in a 4x4 grid, plotting all the four different kinds
