---
title: "Challenge 10: Using purrr package"
author: "Surya Praneeth Reddy Chirasani"
description: ""
date: "01/28/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_10
  - debt
---

```{r}
#| label: setup
#| warning: false
library(tidyverse)
library("readxl")
knitr::opts_chunk$set(echo = TRUE)
```

## Using purrr library to apply a function on multiple columns

I am using debt dataset which I have used in Challenge 9 too, where I have written a function to calculate z_score on a numeric column
```{r}
data <-read_excel("_data/debt_in_trillions.xlsx")
data
```

This is the z_score function I have written in Challenge 9
```{r}
z_score <- function(x){
  stat <- (x - mean(x,na.rm=T))/sd(x,na.rm=T)
  return(stat)
}
```

I am applying map function to calculate z_score for all values of a column in given columns below. The output is a list object with values printed in the same order of the given input columns. 

```{r}
mtcars_stats <- map(list(data$Mortgage, data$`HE Revolving`, data$`Auto Loan`, data$`Credit Card`, data$`Student Loan`,  data$Other, data$Total), z_score)
mtcars_stats
```


